### 2016-11-22 bare_conf.tex
First draft completed

\documentclass[conference]{IEEEtran}

% *** CITATION PACKAGES ***
%
\usepackage{cite}

% *** GRAPHICS RELATED PACKAGES ***
%
\ifCLASSINFOpdf
\usepackage[pdftex]{graphicx}
  % declare the path(s) where your graphic files are
\graphicspath{{./img/}}
  % and their extensions so you won't have to specify these with
  % every instance of \includegraphics
\DeclareGraphicsExtensions{.pdf,.jpeg,.png}
\else

\fi

\usepackage{stfloats}
% *** PDF, URL AND HYPERLINK PACKAGES ***
%
\usepackage{url}

% correct bad hyphenation here
\hyphenation{op-tical net-works semi-conduc-tor}

\begin{document}

% paper title
% Titles are generally capitalized except for words such as a, an, and, as,
% at, but, by, for, in, nor, of, on, or, the, to and up, which are usually
% not capitalized unless they are the first or last word of the title.
% Linebreaks \\ can be used within to get better formatting as desired.
% Do not put math or special symbols in the title.
\title{Bare Demo of IEEEtran.cls\\ for IEEE Conferences}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{\IEEEauthorblockN{Michael Shell}
\IEEEauthorblockA{School of Electrical and\\Computer Engineering\\
Georgia Institute of Technology\\
Atlanta, Georgia 30332--0250\\
Email: http://www.michaelshell.org/contact.html}
\and
\IEEEauthorblockN{Homer Simpson}
\IEEEauthorblockA{Twentieth Century Fox\\
Springfield, USA\\
Email: homer@thesimpsons.com}
\and
\IEEEauthorblockN{James Kirk\\ and Montgomery Scott}
\IEEEauthorblockA{Starfleet Academy\\
San Francisco, California 96678--2391\\
Telephone: (800) 555--1212\\
Fax: (888) 555--1212}}

% use for special paper notices
%\IEEEspecialpapernotice{(Invited Paper)}

% make the title area 
\maketitle
\begin{abstract} This work is focused on the prediction of affective
states using non-obtrusive sensors. Non-obtrusive sensors are preferred in this
work because of a hypothesis that states that an obtrusive sensor would alter
the affective states of a user. Specifically, the method uses data gathered from
a keyboard and mouse dynamics plugin embedded in a tutoring system designed for
the teaching of computer programming in the Python language. The tutoring system
requires from the user to practice their programming skills in order to progress
through the course, and the user’s creativity plays an important role in the
solving of programming exercises. Also a pre-processing method is applied, which
generates feature vectors consisting of 39 features representing each
interaction of the users by recording their keyboard and mouse dynamics. 149
feature vectors were generated and used as input of four different
classification algorithms: Naïve Bayes, J-48, k-NN, and Artificial Neural
Networks. These classification models are then compared by their performances
when classifying learners into five different affective states: boredom,
frustration, distraction, relaxation and engagement. Each of the classification
models achieved an accuracy of around 75\% and kappa of around 0.45, and, in
average, J-48 was the better algorithm with an accuracy of 75.97\%. Results show
that the data gathered from the non-obtrusive sensors can successfully be used
as another input to classification models in order to predict an individual's
affective states during their interaction with a programming tutoring system.
\end{abstract}

% no keywords

\section{Introduction} 
Introductory programming courses are
generally regarded as difficult \cite{robins2003learning,lahtinen2005study} 
and there is a common conception that they often have a high
failure rate \cite{bennedsen2007failure}. There are multiple factors involved.
Jenkins \cite{jenkins2001motivation} argues that some are deeply related with the expectations,
attitudes, and previous experiences of the teaching staff and students. Another
factor is the nature of the subject that involves the learning of new abstract
constructs, syntax and tools. Also groups are often heterogeneous and thus it is
difficult to design courses that are beneficial for everyone. Many students
begin to program when they are at their first year of university and until then
are confronted with a totally new topic that does not respond to their habitual
study approaches. Dijkstra \cite{dijkstra1989cruelty} argues that the subject of programming is very
problem-solving intensive, and it requires high precision because even the
slightest perturbation can render a program totally worthless. These
difficulties often frustrate students. 
Jenkins \cite{jenkins2001motivation,jenkins2002difficulty} notes that many
students expect the course to be difficult and come with the notion that they
will have to struggle, others may have a stereotyped image of a programmer,
these beliefs can negatively affect their initial motivation. During the course,
novice programmers experience many emotions, for instance frustration and
confusion when they cannot find a bug in a program, but also joy when they
successfully run a challenging program for the first time. They can also become
bored if they found the exercises too repetitive or too easy. Learning to
program is a difficult task, where emotions play a significant role. Research
has gone far to understand the role of emotion in learning, both in the process
and the outcome; for instance, in e-learning  \cite{kort2001affective,rossin2009effects}
and in programming  \cite{rodrigo2009affective,jenkins2001motivation,
bosch2013emotions,khan2007mood}.

It is expected that flow/engagement is the ideal affective state in which
students tend to be most capable of acquiring meaningful information through the
learning process. Flow is defined by Csikszentmihalyi \cite{csikszentmihalyi1990flow}
as a mental state in which a person performing an activity is fully immersed in a feeling of
energized focus, full involvement, and enjoyment. Engagement is also a positive
affect where students are thought to be more involved behaviourally,
intellectually, and emotionally in their learning tasks \cite{bangert2002teacher}.
When designing instruction, an objective is to assign learning activities
that result challenging to students, but not much as to frustrate them. For
this, experienced instructors gauge a student’s affective state and then assign
an activity with the appropriate level of difficulty. In affect-aware
intelligent learning environments (ILEs), emotions have a central role in the
students' user model. However, recognition of emotions is a difficult task, even
for humans. Instructors use their social intelligence often to recognize
students’ affective states. In class a good instructor habitually reads the
faces of students in the classroom to see if they are confused, bored or
engaged; then decides what to do next. ILEs can be enhanced if they can offer an
automatic perception of emotions. The recognition and simulation of human
affects are becoming important fields of study, as many researchers have
demonstrated affect-aware computers can provide better performance in assisting
humans \cite{picard2001toward}. When ILEs embrace these methodologies and techniques a
problem arises. In order to perform recognition of the users' affective states,
sensors must be used to gather data. Some sensors rely on physiological readings
and must be in physical contact with students. These sensors can be considered
intrusive or invasive, and can disrupt the student’s learning experience 
\cite{zhai2008stress,sidney2005integrating,arroyo2009emotion}.
Other sensors such as video cameras can also be considered invasive, since they always transmit the
identity, appearance and behaviour along with emotional data \cite{picard2001toward}.

In this work, a method for the recognition of affective states through keystroke
and mouse dynamics is proposed. The hypothesis is that students vary the dynamic
of keystrokes according to their affective state when programming.  The method
has three main components: a Javascript library to capture keyboard and mouse
dynamics from a browser-based editor, a pre-processing step whose output is a
feature vector that is sent to the third component, a classification algorithm.
An experiment was conducted where students solved a series of programming
exercises using a web based learning environment. Using only the data from the
student’s keyboard and mouse dynamics six affective states where recognized for
each of the attempts at solving the exercises. Results obtained from the
experiment are promising. The affective states recognized include the most
common in novice programmers according to the study of Bosch, D'Mello \& Mills
\cite{bixler2013detecting}: boredom, engagement/flow, confusion and frustration. For this experiment
four binary classifiers were compared: J-48 decision trees, k-nearest neighbour
classifier, feed forward neural network, and a naïve Bayes. The output of each
classifier determined if the student experienced the affective state during the
exercise. The proposed method could be used in an ensemble with other sensor
channels to improve the non-invasive recognition of the students’ affective
states when they are working on a programming task. In order to determine what
affective states a student was experiencing, an Experience Sampling Method (ESM)
was used by Csikszentmihalyi \& Larson \cite{kubey1996experience}. After the students successfully
solve a programming exercise, they are presented with an ESM survey that asks
what they were feeling during their solving of the exercise.

If a relationship could be modelled, ILEs could adapt the instruction using the
same devices required to input the program to the computer. This could be
important because writing programs in a computer is an important learning
activity when learning programming. In study by Lahtinen, Ala-Mutka \& Järvinen
\cite{lahtinen2005study} is reported that students rated ``working alone on programming
coursework'' as a more useful situation than lectures.

The structure of this work is organized as follows: Section Related Work
presents a series of works related to the proposed method in this paper; Section
Proposed Method describes the proposed method for the recognition of affective
states in a web based learning environment for the teaching of programming
languages; Section Experiment explains the experimental evaluation of the
proposed method, following by a Results sections and finally, a Conclusion and
Future Work are discussed.

\section{Related Work} 
Affect recognition is an active field of research. Many
methods have been proposed, some require the intervention of the user to fill up
questionnaires or forms; these selections remain static until the user changes
the values. These methods are easy to implement, but cannot detect dynamic
changes.  A more dynamic approach requires the use of sensors to capture
affective states as they change. The context, the environment and the learning
activity determine what kind of sensors can be used. The most common learning
environment is the classroom, a physical space with context to facilitate
learning. But learning is possible in a wide variety of settings, such as
outside-of-school locations and outdoor environments these are sometimes
referred as ubiquitous learning environments, an example of such environments is
the work of Yang \cite{yang2006context} in a context-aware environment, but missing affective
information. There are also virtual learning environments such as the one proposed by Dillenburg
\cite{dillenbourg2002virtual}  where learners can have avatars, and virtual places
where they can play roles and socialize.

In the general context of learning there have been some approaches for affect
recognition. The work of Kapoor y Picard \cite{kapoor2005multimodal} uses a multi modal
approach using sensory information from facial expressions and postural shifts
of the learner combined with information about the learner's activity on the
computer; the learning activity was solving a puzzle on a computer. They report
using multimodal Gaussian Process approach achieving an accuracy of over 86\%.
Sidney, et al., \cite{sidney2005integrating} enhances the intelligent tutor system AutoTutor with
affective recognition using nonintrusive sensory information from facial
expressions, gross body movements, and conversational cues from logs.  The
subject material of that system consisted in lectures of computer literacy.

Elliott, Rickel y Lester \cite{elliott1999lifelike,d2008autotutor} propose the integration of an affective reasoner
with an agent in a virtual environment. In this work an agent called Steve
responds emotionally to other agents and interactive users. The agent simulates
his emotions through different multimedia modes including facial expressions and
speech. Affective reasoning agents were used for training, by putting students
in work related situations. Agents could react to the behavior of students, for
instance if a student was being careless in a task and was in a dangerous
situation, Steve would show distress or fear. Also in a virtual environment the
work of McQuiggan, Robison \& Lester \cite{mcquiggan2010affective} extends this line of research by
investigating the affective transitions that occur throughout narrative-centered
learning experiences. The analysis of affective state transitions in this work
replicated the findings by D’Mello et al. \cite{d2008autotutor} and Baker et al. 
\cite{rodrigo2009affective} where
also engagement/flow dominated self-reported affect. For outdoor environments
Shen, Wang \& Shen \cite{shen2009affective} augment a pervasive e-learning platform with affective
recognition.  The results about emotion recognition from physiological signals
achieved a best-case accuracy (86.3\%) for four types of learning emotions.

Bosch, D'Mello \& Mills \cite{bosch2013emotions} analysed the relationship between affective states
and performance of novice programmers when they were learning the basics of
computer programming in the Python language. The results of their study
indicated that the more common emotions students experienced were engaged
(23\%), confusion (22\%), frustration (14\%), and boredom (12\%). It was useful
to consider these results, as it presented evidence of what affective states
need to be targeted in order to obtain less biased data. For example, if a less
common emotion was chosen, a classifier could opt to classify any feature vector
as not experiencing such emotion. Nevertheless, the classifier would obtain
accurate results, although the classifier would be inaccurate at determining if
a feature vector was actually experiencing the given affective state. Similar to
the previous work, Rodrigo et al., \cite{rodrigo2009affective} observed which affective states and
behaviours relate to student's achievement within a basic computer science
course. The authors found that confusion, boredom and engagement in IDE-related
(on-task) conversation are associated with lower achievement.

There has been very little research reported on the effectiveness of the use of
keyboard and mouse dynamics as a sensory channel for affective recognition, and
the few have not been focused on programming. The preliminary work by Zimmermann
et al. \cite{zimmermann2003affective} describes a method to correlate user’s interactions (keyboard and
mouse) with an emotional state. Several physiological parameters were measured
concurrent with the task. The parameters included respiration, pulse, skin
conductance level, and corrugator activity. The task assigned to subjects was to
shop on an e-commerce website for office-supplies and finally write a predefined
message to the shop operator. Subjects where watching videos in order to change
their emotional state during the experiment. The method was validated using
self-reported emotions using the self-assessment-manikin (SAM), devised by Lang
\cite{lang1980behavioral}, designed to assess the dimensions valence, arousal and dominance
directly by means of three sets of graphical manikins. Results show that the
film clips were effective in inducing the expected mood changes, but no further
empirical results are presented. The work of Vizer, Zhou \& Sears \cite{vizer2009automated} also
uses sensory data based on the time elapsed between each key press, the task
given to users was to write a free-text and used linguistic features in order to
recognize both physical and emotional stress. The results show a classification
accuracy of 62.5\% for physical stress and 75\% from emotional stress; authors
argue that these results are comparable with other approaches in affective
computing. They also stress that their methods must be validated further in
other contexts. Moods may have an impact on programmer’s performance according
to Khan, Hierons y Brinkman \cite{khan2007mood}. It may be possible to detect moods
on the basis of information regarding the programmer’s use of the keyboard and
mouse, and to integrate them into development environments that can improve
programmer performance. They briefly describe a further experiment that could
use keyboard and mouse but only as future work. There are other studies about
the behaviour of programmers, which are not directly concerned with affective
states but are nevertheless important to their performance. Eye tracking in
computing education is proposed in the work Busjahn et al., \cite{busjahn2014eye}, and it is
also used for assessing learner´s comprehension of C++ and Python programs in
Turner et al., \cite{turner2014eye}. Blikstein \cite{blikstein2011using} proposed 
an automated technique to
assess, analyse and visualize students learning computer programming. Blikstein
employs different quantitative techniques to extract students’ behaviours and
categorize them in terms of programming experience.

\begin{figure}[!t] 
\centering 
\includegraphics[width=2.5in]{KMDAffective.png} 
\caption{Code Execution and Affect Recognition Architecture.}
\label{fig_process} 
\end{figure}

Research works related to Keystroke Dynamics (KD) are carried out either using
fixed-texts, or free-texts Gunetti y Picardi \cite{gunetti2005keystroke}.
KD performed on fixed-texts
involves the recognition of typing patterns when typing a pre-established fixed-
length text, e.g., a password. In the other case, free-text KD achieves the
recognition of typing patterns when typing a text of arbitrary-length, e.g., a
description of an item. However, as noted by Janakiraman y Sim \cite{janakiraman2007keystroke},
most of
the research regarding KD is done on fixed-text input, the reason being that
fixed-text KD usually yields better results than free-text KD. Yet, the authors
of this work share the opinion with Janakiraman, R., and Sim, T., that it would
be more useful if KD can handle free text as well as fixed text, this is also a
requirement if the text is a program.

Although the use of Keystroke Dynamics (KD) is found in several research works
as a biometric measure, its use for identifying affective states is rare in
comparison. Epp, Lippold \& Mandryk \cite{epp2011identifying} effectively used KD in conjunction
with decision-tree classifiers for the identification of 15 affective states.
Although their work was based on fixed-text their technique to extract a feature
vector was an inspiration for the proposed method in this work. As for free-text
KD, Bixler y D'Mello \cite{bixler2013detecting} present a method for the identification of boredom
and engagement based on several classification models.

Regarding Mouse Dynamics (MD), some research has been conducted for the
identification of affective states, although, as with the case of KD, MD is
mainly used as a biometric measure for authentication. Salmeron-Majadas, Santos
and Boticario \cite{salmeron2014exploring} use both MD and KD to predict four affective states using
five different classification algorithms. Bakhtiyari y Husain \cite{bakhtiyari2014fuzzy} discuss a
method based on fuzzy models for the recognition of emotions through KD, MD and
touch-screen interactions. For a broad review of emotion recognition methods
based on KD and MD, the work by Kolakowska \cite{kolakowska2013review} is recommended.

\section{Proposed Method} 
The goal of this work is to propose an affective
recognition method based on the sensory data provided by the keyboard and mouse
dynamics generated by a learner as he/she types a programing exercise. The main
components of the method are depicted in Figure \ref{fig_process} A brief explanation of the
whole process is explained next, details come later. The process starts when the
learner begins to type a program in a browser-based editor. As she types or
moves the mouse the dynamics are recorded. When the learner submits the code to
evaluation, the request includes the sensory data along with the code and
information about the session. In the server, the code is evaluated in an
external virtual machine that provides a sand box to prevent malicious or
erroneous code to halt the server. When the result is ready, it is recorded
along with the sensory data and sent to the preprocessing module. The output is
a feature vector, ready for classification. A previously trained classifier is
responsible for the classification, which outputs the predicted affective state.
The method does not consider other sensory data, but it could be integrated with
other sensory inputs in a multi-modal approach.  Each of the steps is explained
in detail next.

\begin{table*}[!t]
\centering
\caption{Table 1. The components of the feature vector; adapted from Epp, Lippold \& Mandryk \cite{epp2011identifying}.}
    \begin{tabular}{ | l | l || l | l | }
    \hline
    ID          & Description           & ID               & Description  \\
    \hline
    1 & Mean duration between 1st and 2nd down keys of the digraphs.  & 21 & Mean duration of the 2nd key of the trigraphs.\\
    \hline
    2 & Standard deviation of the previous feature.                   & 22 & Standard deviation of the previous feature.\\
    \hline
    3 & Mean duration of the 1st key of the digraphs.                 & 23 & Mean duration between 2nd key up and next key down of trigraphs.\\
    \hline
    4 & Standard deviation of the previous feature.                   & 24 & Standard deviation of the previous feature.\\
    \hline
    5 & Mean duration between 1st key up and next key down of the digraphs. & 25 & Mean duration of the third key of the trigraphs.\\
    \hline
    6 & Standard deviation of the previous feature. & 26 & Standard deviation of the previous feature.\\
    \hline
    7 & Mean duration of the 2nd key of the digraphs. & 27 & Mean duration of the trigraphs from 1st key down to last key up.\\
    \hline
    8 & Standard deviation of the previous feature. & 28 & Standard deviation of the previous feature.\\
    \hline
    9 & Mean duration of the digraphs from 1st key down to last key up. & 29 & Mean number of key events that were part of the graph.\\
    \hline
    10 & Standard deviation of the previous feature. & 30 & Standard deviation of the previous feature.\\
    \hline
    11 & Mean number of key events that were part of the graph. & 31 & Mean duration of mouse key presses.\\
    \hline
    12 & Standard deviation of the previous feature. & 32 & Standard deviation of the previous feature.\\
    \hline
    13 & Mean duration between 1st and 2nd down keys of the trigraphs. & 33 & Mean duration of all mouse movements.\\
    \hline
    14 & Standard deviation of the previous feature. & 34 & Standard deviation of the previous feature.\\
   \hline
    15 & Mean duration of the 1st key of the trigraphs. & 35 & Mean distance in pixels in the X coordinate. \\
    \hline
    16 & Standard deviation of the previous feature. & 36 & Standard deviation of the previous feature.\\
    \hline
    17 & Mean duration between 1st key up and next key down of trigraphs. & 37 & Mean distance in pixels in the Y coordinate.\\
    \hline
    18 & Standard deviation of the previous feature. & 38 & Standard deviation of the previous feature.\\
    \hline
    19 & Mean duration between 2nd and 3rd down keys of the trigraphs. & 39 & Number of attempts\\
    \hline
    20 & Standard deviation of the previous feature. &   &  \\
    \hline
  \end{tabular}
\end{table*}

\subsubsection{Capturing the Keystroke and Mouse Data} 
While a student is trying
to solve an exercise, a script coded in JavaScript is running in the background,
which captures every keystroke, mouse movement and mouse button press. Each
capture of these events records a timestamp in milliseconds (using the method
getTime() of Javascript's built-in class Date) that describes when the event
occurred. If the event is a keystroke, the script captures what key was
specifically pressed, and what type of event occurred, it can be either a key-
down or a key-up event. If it is an event related to a mouse button press, the
key code of that button is recorded, as well as the type of event occurred again
key-down or key-up. Finally, if the event was a mouse movement, the mouse
coordinates inside of the web browser are recorded. The script monitors the
mouse position every 100 milliseconds, and only if the position has changed, it
records the new position. Each time a learner tries to evaluate a program, all
the data generated is sent to the server along with the code. When the result of
the execution is returned, all records are cleared and the process starts again.
There is no problem if the user leaves for a long period of time, because no
event will be triggered. If a user copies and then pastes the code from another
source, this will be recorded. There is a limitation, only the browser-editor
must be used; this could be a problem for more advanced programmers needing
specialized editors.  On the other hand a browser-based editor with the
corresponding remote execution, does not require the installation of
interpreters or compilers in learner’s computer. The code could even be written
in a mobile device or any web-enabled device.  The interface of the web-based
editor is shown in Figure \ref{fig_editor}. Programming exercises are evaluated using unit
tests; the results of the evaluation are shown to users. If all tests the
program is considered to be correct. An example execution is shown in the left
side image of Figure \ref{fig_editor}. In this example the learner was asked to write a
function to add two numbers, in Python. The source code for the Protoboard web
based learning environment including the KD and MD functions are open source and
available as Github repositories at http://git.io/vJlUV also the code for the
sandbox is in http://git.io/vJlUj.

\begin{figure}[!t] 
\centering 
\includegraphics[width=2.5in]{editorRresult.png} 
\caption{Web-based editor used for code evaluation.}
\label{fig_editor} 
\end{figure}
\begin{figure}[b!] 
\centering 
\includegraphics[width=2.5in]{KeyDyn.png} 
\caption{Keystroke Dynamics, adapted from Sim \& Janakiraman \cite{sim2007digraphs}.}
\label{fig_KD} 
\end{figure}
\subsubsection{Preprocessing of the Keystroke and Mouse Data} 
The raw data
obtained from the script needs to be preprocessed to obtain a feature vector.
Basically, this pre-processing consists in measuring the delays between key-
down, key-up or mouse-move events triggered during an exercise. These events
have the dynamic shown in Figure 3; for example when typing the word ‘key’ a
user first presses the letter K and triggers the key-down event, this is
indicated with an arrow that changes the state of the key, the time of the event
is important and also recorded. Only the event data is received from the
browser, in order to generate feature vectors, the patterns and rhythm users
have when pressing consecutive keys are captured measuring the delays between
events, as it is a common practice when dealing with keystroke dynamics. In this
work, the definitions proposed by Sim y Janakiraman \cite{sim2007digraphs} are used. Held
time (Ht) is defined as the time between a key-down and a key-up of the same
key, this would be Ht(K) in the example. Inter key time (It) is defined as the
time between two consecutive key-down events, this time could be negative is the
second key is pressed before the first is released. A sequence is defined as a
list of consecutive keystrokes. In the above example valid sequences could be
[‘K’, ‘E’], [‘E’, ‘Y’] and [‘K’, ‘E’, ‘Y’], the first two are digraphs and the
third is a trigraph. While there can be sequences of any size in a text, working
only with digraphs and trigraphs is preferred. Even if only digraphs and
trigraphs are considered, the amount found in free-text is too large and not
very useful for a feature vector as noted by Epp, Lippold y Mandryk (2007) so
they propose to use statistical measures to capture the patterns; the same
strategy is used in this work. The averages and standard deviations are
calculated from the delays between the events of the digraphs and trigraphs in
each program.

To calculate the average and standard deviations of these presses, the delays
between a key-down and a key-up event of the left button clicks are used. In
addition to these averages and standard deviations of the delays between
keystrokes and mouse button presses, the average and standard deviations of the
number of total events contained in a digraph and a trigraph are calculated.
These features are proposed and explained by Epp, Lippold y Mandryk \cite{epp2011identifying}. Most
of the times, a digraph should contain four events, while a trigraph six.
However, sometimes an individual can start a digraph or a trigraph before ending
the previous one. These additional features represent these particular cases,
and could be meaningful for the estimation of a learner's affective states.
Regarding the mouse movements, the average and standard deviation of the
duration of each mouse movement, and the averages and standard deviations of the
movements in the axes X and Y are also calculated. Lastly, a final feature is
added to preprocessing of the data. The web tutorial recorded how many attempts
a student required before successfully solving an exercise. This number of
attempts is also included in the feature vector. The final feature vector
consists of 39 features; these are based on the work of Epp, Lippold \& Mandryk
\cite{epp2011identifying} and are shown in Table 1. (The code used for this step is available at:
https://github.com/amherag/keyboard-mouse-dynamics).

Once feature vectors are obtained from an experiment, the generated dataset is
normally used as training data for a classifier. Researchers of affective
recognition have used a wide variety of classification algorithms. As a proof of
concept four well known classification algorithms where compared: k-Nearest
Neighbors (k-NN) algorithm, a feed forward neural network trained with back-
propagation, a naïve Bayes classifier and finally a decision trees algorithm for
rational data (J-48). Details for these algorithms, can be found in the textbook
by Tan, Steinbach, \& Kumar \cite{tan2006introduction}. The accuracies obtained by these algorithms
as well as the parameters used are reported in the next section.

\section{Experiment} 
The aim of this research is to evaluate the use of keyboard
and mouse dynamics as an appropriate sensory input for an affective recognition
system. The context of use is a learning environment for programmers. In
particular for learning activities consisting in writing short programs
interactively. An experimental approach was adopted with this aim. Sensory and
quantitative data was collected from learners as they were enrolled in a basic
course of Python programming. This data was then pre-processed using the method
described earlier, and together with the quantitative data obtained from users,
classifiers were trained and validated. The results were then compared and came
to conclusions regarding the effectiveness of the method. The goal given to
subjects was to solve as many exercises as they could in a period of two weeks.
There was neither time limit nor a minimum amount of time required for a
participant while trying to solve the exercises or complete the tutorial. The
participants were able to stop and resume their interaction with the system at
any time.

\begin{figure}[!t] 
\centering 
\includegraphics[width=2.5in]{Completed.png} 
\caption{Completed activities and exercises and by learner in ascending order.}
\label{fig_completed} 
\end{figure}

\begin{figure}[!t] 
\centering 
\includegraphics[width=2.5in]{classDist.png} 
\caption{Class distribution of emotions as reported by learners.}
\label{fig_class} 
\end{figure}

A tutorial was developed to
obtain the necessary data using Protoboard, a web-based programming learning
environment (the latest version can be found online at
http://python.protoboard.org/). Protoboard has the functionality described in
the proposed method and shown schematically in Figure 1. Users-log in to
Protoboard using their account credentials from a popular social network
(Facebook).  The web tutorial begins with three introductory videos that explain
the fundamentals of programming in Python, and how to solve the programming
exercises in the course. What follows after these videos, are 13 programming
exercises that the students need to solve in consecutively. Exercises are of
incremental difficulty. In Figures 2 and 3, the interface of this platform is
presented. For this experiment Protoboard was configured to allow unlimited
tries to solve an exercise.

A total of 55 volunteers, with no previous experience in Python, where recruited
as a response to three announcements in a special interest group of programming
students in a social network. Learners were required to login with their
Facebook credentials, this had the advantage of not requiring the system to
maintain users’ credentials or profile information as this is kept outside in an
external service. Unfortunately this decision also kept some prospects from
volunteering; a subject reported that he normally creates temporary email
accounts to try new web services. Other requirement was that they where all
adults, this age is 18 years in Mexico where the experiment was conducted. The
requirement of the social network account stopped several prospects from
volunteering, but on the other hand gave researchers access to public
information about the subjects. All the learners were either current students or
graduates of Computer Systems Engineering. Their ages were in the range of 18 to
30 years old. Although no experience in software programming was needed, as the
web tutorial's course is of a very basic level, all the participants were
required had completed at least one course in software programming. Results from
the initial survey are shown in Figure 4.
\begin{table*}[!t]
\centering
\caption{The performance of the classifiers.}
    \begin{tabular}{ | c | l | l | l | l | }
    \hline
    Affect          & Naïve Bayes           & J-48                & k-NN              & ANN \\
    \hline
    Flow/engaged    & 79.00\% (0.11) 0.50   & 80.48\% (0.10) 0.51 & 76.14(0.09) 0.43    & 78.43(0.12) 0.45 \\
    \hline
    Relaxation      & 71.10\% (0.09) 0.40   & 72.57\% (0.11) 0.45 & 69.14\%(0.08)  0.37 & 71.24\%(0.07)  0.34\\
    \hline
    Distraction     & 70.33\% (0.13)  0.43  & 71.00\% (0.10) 0.43 & 74.00\%(0.06)  0.50 & 72.52\%(0.07)  0.39\\
    \hline
    Frustration     & 74.71\% (0.09) 0.49   & 73.86\% (0.12) 0.45 & 72.62\%(0.09)  0.42 & 78.0\%(0.09)  0.51 \\
    \hline
    Boredom         & 74.71\% (0.13) 0.47   & 84.57\% (0.06) 0.59 & 83.81\%(0.06)  0.58 & 76.19\%(0.09)  0.45\\
    \hline
    \end{tabular}
\end{table*}

In order to determine what affective states a student was experiencing, an
Experience Sampling Method (ESM) was used (Csikszentmihalyi y Larson, 1987).
After the students successfully solve a programming exercise, they are presented
with an ESM survey that asks what they were feeling during their solving of the
exercise. A very brief description is given about what to do in this survey,
followed by statements the students need to answer according to how they were
feeling. As an example, the statement “I was feeling frustrated” is presented,
and a student needs to answer either “Strongly agree,” “Agree,” “Neutral,”
“Disagree,” and “Strongly Disagree.”

After the two weeks of the experiment, only four learners completed all of the
programming exercises and 22 did not completed any. Out of the total activities
available (videos, survey and questionnaires) only two completed all. Figure 5
shows the number of exercises and activities completed by each learner.

The participants' interaction generated a total of 142 feature vectors one for each
successfully completed exercise. The affective states reported by learners after
completing each exercise was grouped in three classes: ‘Yes’, ‘Neutral’ and
‘No’. The class distribution of the emotions reported is shown in Figure 6. This
results show that the most common emotions were flow/engagement (72\%) and
relaxation (61\%) while few learners reported distraction (5\%), frustration
(8\%) and boredom (2\%). This distribution is different from what was reported
by D’Mello et al. \cite{bixler2013detecting} and Rodrigo et al. \cite{rodrigo2009affective}.
Although Flow/Engagement was
also the predominant class, the distribution is skewed to the first two. There
are some possible reasons for this; first the majority of students had
experience in programming, so learning new one is not that problematic, a second
reason could be the freedom users had to abandon the activities, perhaps
frustrated or bored learners simply quit the tutorial. A high number of learners
(49\%) did not completed any exercise and the once who completed more where also
the more experienced.

Classifiers where implemented using the software Rapid Miner using a feature
selection module as part of the process. A genetic algorithm was used in the
user selection module. In the end, the subset for the frustration classifier
consists of 11 features, and the subset for the boredom classifier consists of
13 features. The performance of the classifiers is shown in Table 1.

The results obtained with this method are satisfactory, considering a free-text
was used. The accuracies and kappa coefficients obtained are close to what is
usually obtained in fixed-text methods, for example, the results presented in
Epp, Lippold y Mandryk (2007). In the case of the kappa coefficients, it is
usual to see values below 0.2 in methods involving free-text. In this case, some
values of kappa were close to 0.5, is important to consider the values of kappa
in these results because the class distribution is not uniformly distributed. As
it is observed in many works in affective recognition, decision trees normally
produce competitive accuracy. In this case the J-48 classifier obtained an
accuracy of 80.48\% marginally better than the others, and an accepted kappa
statistic for the kind of problem. Artificial neural network give the highest
accuracy in the classification of frustration.

\section{Conclusion}
The aim of this research was to evaluate the use of keyboard and mouse dynamics 
as an appropriate sensory input for an affective recognition system. The context
of use would be learning environment for programmers. In particular for learning
activities consisting in writing short programs interactively. An experimental
approach was adopted with this aim. The proposed method in this work obtained
satisfactory results. It is usual for a classification method based on free-text
to obtain an accuracy and kappa below their counterparts of fixed-text, and in this
work the results obtained were similar to those works applied to fixed-text 
dynamics. A possible explanation of this would be the addition of the mouse 
dynamics features and the additional preprocessing performed on the feature vectors. 
While these are promising results further experiments are needed. Other experiments
should focus on novice programmers.

Future work will focus on extracting other features from the learners’ interaction
like behavioral data, as proposed by Blikstein (2011); the data includes the size
of the program and changes in the code. Other authors also propose treating the use
of certain keys as features e.g., the delete key because is manly used for correcting
sentences. Also a multimodal approach is needed.

% use section* for acknowledgment 
\section*{Acknowledgment} 
This received funding from Tecnólogico Nacional de México through the project: Técnicas de
computación inteligente para el secuenciado adaptativo de ejercicios de
programación en la nube.

% trigger a \newpage just before the given reference 
% number - used to balance the columns on the last page 
% adjust value as needed - may need to be readjusted if 
% the document is modified later %\IEEEtriggeratref{8} 
% The "triggered" command can be changed if desired:
%\IEEEtriggercmd{\enlargethispage{-5in}}

% references section

\bibliographystyle{IEEEtran} 
% argument is your BibTeX string definitions and bibliography database(s) 
\bibliography{programming} 
\end{document}